Paper 1: Triangulating trends in inequality around the world from secondary databases

### 2023-04-19 PhD_pages/Paper_1/data_appendix/1_H_figures_and_tables.qmd
Paper 1: Further drafting and adding figures

```{python}
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share'],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1
                ]
            }

@@

## Tables

#### Table 1

@@

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary['PIP: Gini']

dfi.export(df_summary,"table_test.png")

```
![](table_test.png)

>>>>

```{python}
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share'],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1
                ], 
                'relative_tolerance': [
                    0.05
                ]
            }

@@

## SECTION 2:

### WID top 1% vs PIP Gini

#### Since 1993

This figure compares relative change in the 'typical' metrics for WID and PIP: Top 1% share (pretax) and Gini
```{python}
# Set-up to compare WID and PIP top 10% share, all countries
reference_vals = [1993, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'WID: Top 1pc share – pretax',
        'PIP: Gini'
        ]

@@

df_prepped = prep_data(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup
    )

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      absolute_or_relative_tolerance = "absolute",
      region_col = 'region'
      )

# df_summary = df_summary['WID: Top 1pc share – pretax']
df_summary = df_summary['PIP: Gini']

df_summary
```

### 2023-04-19 PhD_pages/Paper_1/paper1.qmd
Paper 1: Further drafting and adding figures

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np
from IPython.display import display, HTML, IFrame, Markdown
import dataframe_image as dfi
import imgkit

from wrangling_and_plotting_functions import *

pd.set_option('display.max_rows', None)

@@

# This specifies the +/- for the 'stable' category in the tables below
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share'],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1
                ]
            }

@@

Two sources of data have been particularly impactful in this regard: the World Bank's harmonized collection of household survey data, published online through its [*Poverty and Inequality Platform (PIP)*](https://pip.worldbank.org/); and the collection of inequality estimates – most notably top income shares – brought together at the [*World Inequality Database (WID)*](https://wid.world/).^[EDIT: PIP used to be Povcal. It draws upon the work of various regional teams. WID used to be WTID.]

These two sources are the only regularly-updated databases providing global coverage on a relatively consistent basis.^[EDIT: Both datasets suffer important limitations with respect to comparability acrosss time and location. These will be discussed in the data section below. Both datasets, however, take significant steps to harmonize, *ex-post*, the available underlying data. In this respect they can be distinguished from databases such as WIID and SWIID [@atkinson_promise_2001]] Both have played substantial agenda-setting roles in academic research, policy and public debate.

Given its extensive coverage of low- and middle-income countries, the World Bank data occupies a central place in international development policy and research. Most notably, the data is relied on by the UN to set and monitor progress towards many of the targets included within Sustainable Development Goals. It has also to date underpinned the majority of academic studies on inequality that take a global view.^[[Including @Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk] TODO: Check and more refs].

The *WID* database emerged as institutional focal-point to host and coordinate inequality estimates produced as part of the burgeoning literature on top incomes [@atkinson_top_2014; @piketty_capital_2014; @alvaredo_top_2013]. This development of this data has opened new avenues for the study of inequality right across the social sciences (TODO: example refs), underpinning a significant re-evaluation of its causes and consequences [@alvaredo_top_2013]. It has also surely had a substantial impact on the public's conception of inequality, contributing to the momentum generated by the Occupy movement and its framing of inequality in terms of 'the 1%' vs the '99%''.

#### Two worlds of inequality measurement

The two . They in some sense represent *two different worlds* of inequality measurement.

In terms of methods, source, conception of inequality and research questions. "The 1 %" – pretax, with many assumptions needed to get to national income. These come as a bundle. Although both institutions continue to take many steps to provide transparency and reproducibility to their estimates, they remain a black box.

Also in the research questions the datasets aim at. Capital vs labour share. .– including fundamental laws of capitalism with regard to capital accumulation (Piketty) –

Most crucially, the worlds also diverge in terms of headline messages as to inequality around the world. Contradictory messages emerge concerning the most basic descriptions from the two databases.

@@

Secondly, the two worlds differ in their conception of global inequality....  decompositions of global inequality into between- and within-country components. (Explain

Incredible efforts, involving the work of hundreds of researchers. Both datasets have continued to improve adn expand. In terms of data availability we are thus in a better position than ever to take a global viw on inequlity. And yet, amidst this wealth of data, limited understanding of very basics concerning inequality around the world.

@@

Many methodolodical choices concerning underlying data sources, the constraints on temporal coverage this gives rise to,  adjust, conception of inequality. These to some extent come bundled together when making the choice between PIP and WID.

In many particular contexts There are known strengths and weaknesses. However in using the datasets a global view on inequality, requires choosing between methodological 'bundles' whose relative strengths and weaknesses, applied globally, are not well understood.

### Connecting the two worlds
*(In this section I will explain the plan for the paper)*

What we can do is triangulate

In this paper, I focus on trends in within-country inequality in recent decades.(In a companion paper, I look at the global distribution.) I line up what I can. I triangulate what we know – What is true irrespective of which 'world' we consider. And Identify what we don't know – the remaining differences – and try to pinpoint where the differences lie.

I do so with respect to two separate time periods, chosen to tradeoffs with respect to global coverage. The first period – from the fall of the Berlin wall. Driven primarily to the steep drop off in availability before this in household survey data collected in PIP.

This maximises global coverage – 100 countries covering roughly 90% of the world poplation. The downside is that this period misses out – transition, liberalisation and structural adjustment. Question begging, by avoiding specifically the period where inequality thought to rise.

To address this, I analyse a second, longer period extending back to roughly 1980.

Explain approach. (Including LIS – which is close to PIP world (indeed PIP takes data from LIS for a number of high income countries.))

(SUMMAR OF FINDINGS)

## Methods and data

### Concepts of global inequality

Adapted three concepts.

### Data

#### PIP

#### WID

#### LIS

Will be used for income concept. i.e. it represents the PIP world – indeed LIS is the source for PIP for a number of high income countries. But the access to microdata allows us to ask more questions for a subset of countries.

#### Coverage and lining-up observations

(Coverage chart - overlayed 3/5 year)

## Results

### Comparison of typical indicators

- Relative change WID 1% and PIP Gini

### Aligning indices

## Comparing inequality trends

The focus of this paper is on trends. But worth pointing out levels. Inequality is much higher in WID than in PIP.

Again this relates to many things – pretax vs disposable income/consumption. Later we will try lining up income concepts more closely. But for global coverage we have to pull together these 'worlds' that combine many differences.

@@

No reason to expect this given the properties of the two measures. If we take lognormality we see... (MAke a chart in appendix - where we plot Gini against top 10 and top 1% assuming lognormality, and then add actual data for a particular country.) How unlognormal the top is would be another way to judge similarity. Rather than doing this, we bridge by aligning the indices.

- We also do absolute change: a percentage point. This is more intuitive. But somewhat arbitrary – differences in levels affect assessment of differences in trends.

@@

@@

>>>>

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np
# from IPython.display import display, HTML, IFrame, Markdown
# import dataframe_image as dfi
# import imgkit

from data_appendix.G_functions_for_figures_and_tables import *

# pd.set_option('display.max_rows', None)

@@

# This specifies the +/- for the 'stable' category in the tables below
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share'],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1
                ], 
                'relative_tolerance': [
                    0.05
                ]
            }

@@

Two sources of data have been particularly impactful in this regard. One is the World Bank's harmonized collection of household survey data, published online through its [*Poverty and Inequality Platform (PIP)*](https://pip.worldbank.org/). Another is the collection of inequality estimates, most notably top income shares, brought together at the [*World Inequality Database (WID)*](https://wid.world/).^[TODO: PIP used to be Povcal. It draws upon the work of various regional teams. WID used to be WTID.]

These two sources are the only regularly-updated databases that provide global coverage of distributional data on a relatively comparable basis.^[EDIT: Both datasets suffer important limitations with respect to comparability acrosss time and location. These will be discussed in the data section below. Both datasets, however, take significant steps to harmonize, *ex-post*, the available underlying data. In this respect they can be distinguished from databases such as WIID and SWIID [@atkinson_promise_2001]]

Both have played substantial agenda-setting roles in academic research, policy and public debate.

The World Bank data is especially prominent within international development policy, notably being used to monitor many of the UN Sustainable Development Goals. Academic research on global inequality has also largely relied on this data  [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk]. *[TODO: Check and more refs.]*.

The WID database emerged along with the burgeoning literature on top incomes, acting as a central repository and, increasingly, coordinating institution. The 1%

## Two worlds of inequality measurement

As well as being the the two  not just two different databases. They in some sense represent *two different worlds* of inequality measurement.

Different methods. These yield different pictures of the extent of inequality and how it has changed over time. Motivated by different ideas about the salient factors shaping inequality.

### PIP

### WID

@@

Decompositions of global inequality into between- and within-country components.

Capital vs labour share. .– including fundamental laws of capitalism with regard to capital accumulation (Piketty) –

### The disconnect between these datasets devaules them both

@@

Trouble is that these are all bundled up. You must chose. On what basis can you make a choice?

Known problems on both sides.

For PIP – top incomes, and lack of comparability of the survey data – notably mix of incme and consumption data.

For WID – tha many assumptions. An emphasis placed on the incorporation of tax data, but when taking a global view confonte by the fact that such tax data is often nonexistent. Top income corrections based on strong assumptions.

Given the presentation, the extent to which different assumptions, indices, methodological choices affects the data and – is not something that can be read off' the data.

In using the datasets a global view on inequality, requires choosing wholesale between methodological 'bundles' whose relative strengths and weaknesses, applied globally, are not well understood.

Also makes for a very confusing terrain for policymakers, media and the public.

The two datasets are incredible efforts, involving the work of hundreds of researchers. Both datasets have continued to improve adn expand. In terms of data availability we are thus in a better position than ever to take a global viw on inequlity. And yet, amidst this wealth of data, limited understanding of very basics concerning inequality around the world.

### Outline: Connecting the two worlds
*(In this section I will explain the plan for the paper)*

In this paper I aim to bring these two worlds of inequality measurement closer. To Triangulate what they do agree on and identify where they disagree.

The first, most obvious, step is simply to compare the in a systemtatic way. This very basic step is curiously lacking in the literature. We begin by comparing trends in each world based on the inequality indices most typical in each: the Gini coefficient for the PIP data, and the top 1% share of pre-tax income for WID.

Given incomplete coverage in both datasets, we 'line up' estimates to reference years to see change over time. We use two periods, presenting two different tradeoffs with respect to data coverage. The first is 1993-2018?? This maximises global coverage – 100 countries covering roughly 90% of the world poplation (Figure XX). The downside is that this period misses out – transition, liberalisation and structural adjustment. Question begging, by avoiding specifically the period where inequality thought to rise.

The second is 1980-2018. Here we lose coverage, especially in the PIP data (Figure XX).

The two measures selected – Gini coefficient of disposable income/consumption, and the top 1% share of pre-tax income – are where the PIP and WID worlds respectively are at their furthest apart. Yet in terms of broad global trends, there is much that even these very different measures can agree on: - x, - y, - z.

There are big differences though:
- Somewhat more pronounced rise among advanced economies.
- Big divergence in trend some of the most populous countries

The rest of the paper is about understanding where those differences emerge.

- Inequality indices chosen. We align. Compare Top 1%. And Gini.
    - Explain results.

- We look at different income concepts for the small number of coutnries for which such data exists. To this end we also include data from the Luxembourg Income Study. Similar to PIP in that it relies on survey data. But provides access to microdata, allowing for the construction of different welfare aggregates. We focus on the countries with the biggest difference in trend and biggest population. I.e. those that contribute most to the difference in the weighted an unweigted global averages.

    What I won't do is a detailed lining up.
    Although both institutions continue to take many steps to provide transparency and reproducibility to their estimates, they remain a black box. Makes it impossible to do this lining up. 
    In many particular contexts – where detailed lining up is possible – there relative strengths and weaknesses may be clear. (This is what Morllei does. And Nolan). But at the global level the makes this impractical.

    - Summarise: really depends on the country.

----

# 2. Comparing the two worlds

@@

Possibly a footnote: No reason to expect this given the properties of the two measures. If we take lognormality we see... (MAke a chart in appendix - where we plot Gini against top 10 and top 1% assuming lognormality, and then add actual data for a particular country.) How unlognormal the top is would be another way to judge similarity. Rather than doing this, we bridge by aligning the indices.

- We also do absolute change: a percentage point. This is more intuitive. But somewhat arbitrary – differences in levels affect assessment of differences in trends. (Or possibly empahasise this measure if there is little difference to the relative threshold.)

@@

To get an overview: I'll use the unweigted and population weighted average. The first represents change in inequality in the average country.

All broken down by region.

## Inequality since 1993

In @fig-pip-wid-1993-2015-all-obs we plot the level of inequality in 1993 and 2015 against each other, using the inequality measures 'typical' of each dataset.

To help with the interpretation, a band is shown along the 45-degree line. Countries falling within this band saw little change in inequality. Countries falling above it saw a substantive rise in inequality between the two periods. Countries falling below it saw a substantive fall. 'Substantive' here is defined as a change of more than 2 percentage points in the case of the PIP Gini coefficient, and more than 1 percentage point in the case of the WID top 1% share.

From the visual inspection, in conjunction with @tbl-pip-wid-1993-2015-all-obs-averages.  In both sets of data, we see a mix of countries with rising and falling inequality. In the case of WID top 1% share, far more rises than falls. In both we see some convergence – . This is largely driven by regional dynamics.

This is confirmed in summary tables.  and this ...
Tolerance: different thresholds due to different range of data. In the appendix we consider relative thresholds (+/-5%), which makes little difference.

@tbl-pip-wid-1993-2015-all-obs-absolute-count and @tbl-pip-wid-1993-2015-matching-obs-absolute-count do this...

Considering all available observations in each dataset, we see: 
- 
-
- Much bigger rise in the weigted average in WID.

Whilst the number of countries available in the two datasets is very similar, they are not the same countries. In particular, LAC is largely absent from WID. (Footnote about LAC in WID?? Are the global results based on linear interpolation?... An interpolation that contradicts the survey data?) Also fewer Sub-Saharan African countries. In PIP, fewer rich countries in Western Europe and its offshoots.

Importantly, ... are not neutral with respect to.The regions where WID is lacking coverage – LAC and SSA – are regions in which inequality was generally falling, according to the PIP Gini coefficient. The reverse is also the case: Regions that PIP are lacking – high income countries in W. Europe and offshoots – were regions where inequality was generally rising, according to the WID top 1% shares data.

As such, differences in coverage already go some way to explaining the differences between the two datasets. 
@fig-pip-wid-1993-2015-matching-obs and @tbl-pip-wid-1993-2015-matching-obs-averages show the data, restricting the observations to those available in both datasets. This brings the two series closer. Explain. (Coverage falls)

It is worth emphasising the similarities:
-
-
-

Nevertheless, there are clearly differences in what – even after aligning coverage. Are these about the metric, or other methodological differences? WE will see later on.

```{python}
# Set-up to compare WID and PIP top 10% share, all observations
reference_vals = [1993, 2015]
max_dist_from_refs = 3

selected_vars = [
        'PIP: Gini',
        'WID: Top 1pc share – pretax'
        ]

output_dict = prep_data_and_change_summary(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      region_col = 'region',
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup
      )

# Store data for OWID upload – prepared further in appendix below
df_owid_1993_2015 = output_dict['owid_format']

fp = 'data_appendix/figures_and_tables/Compare WID and PIP inequality data 1993 vs 2015 (Joe temp).csv'

df_owid_1993_2015.to_csv(fp, index=False)

#### Figures 1-2

::: {.panel-tabset}
## All available observations

![PIP Gini and WID Top 1% shahre in 1993 vs 21015 – All available observations](images/figures/PIPGini-vs-WIDTop1-1993-vs-2015-01.png){#fig-pip-wid-1993-2015-all-obs}

## Only observations available in both datasets
![PIP Gini and WID Top 1% shahre in 1993 vs 21015 – Matching observations only](images/figures/PIPGini-vs-WIDTop1-1993-vs-2015-02.png){#fig-pip-wid-1993-2015-matching-obs}

#### Tables 1-2
::: {.panel-tabset}
## All available observations

```{python}
#| label: tbl-pip-wid-1993-2015-all-obs-absolute-count
#| tbl-cap: Change in inequality 1993 vs 2015 - All availalbe observations
output_dict['all_sources_False']['absolute_counts']

## Only observations available in both datasets

```{python}
#| label: tbl-pip-wid-1993-2015-matching-obs-absolute-count
#| tbl-cap: Change in inequality 1993 vs 2015 - Matching observations only
output_dict['all_sources_True']['absolute_counts']

#### Tables 3-4

::: {.panel-tabset}
## All available observations

```{python}
#| label: tbl-pip-wid-1993-2015-all-obs-averages
#| tbl-cap: Unweighted and weighted average in 1993 vs 2015 - All availalbe observations
output_dict['all_sources_False']['averages']

## Only observations available in both datasets

```{python}
#| label: tbl-pip-wid-1993-2015-matching-obs-averages
#| tbl-cap: Unweighted and weighted average in 1993 vs 2015 – Matching observations only
output_dict['all_sources_True']['averages']

```{=html}
<!-- Add icon library -->
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">

<a href="data_appendix/figures_and_tables/PIP: Gini vs WID: Top 1pc share – pretax – 1993 to 2015 – allObsFalse.csv" download="Gini_vs_Top1_1993_2015.csv">
<!-- Auto width -->
<button class="btn"><i class="fa fa-download"></i> Download full data</button>
</a>
```

#### Figures 1-4

```{=html}

<html>
<body>

<input type="radio" name="1993-vs-2015_radio" value="PIP" checked="checked" onclick="chartSwitchFunction1993_2015_img()">&nbsp;PIP: Gini coefficient
&nbsp;
&nbsp;
<input type="radio" name="1993-vs-2015_radio" value="WID" onclick="chartSwitchFunction1993_2015_img()">&nbsp;WID: Top 1% share

<br>
<label for="1993-vs-2015_radio_obsCheck">Show only observations available in both datasets:</label> 
<input type="checkbox" id="1993-vs-2015_radio_obsCheck" onclick="chartSwitchFunction1993_2015_img()">

<img src="images/figures/pip-gini-1993-vs-2015-all-obs.png" id="1993-vs-2015_plot" width="500">
<br>

</body>
<script>
function chartSwitchFunction1993_2015_img() {

  var ele = document.getElementsByName('1993-vs-2015_radio');

  var checkBox = document.getElementById("1993-vs-2015_radio_obsCheck");

  if (checkBox.checked == true){
    if(ele[0].checked)
        document.getElementById("1993-vs-2015_plot").src = "images/figures/pip-gini-1993-vs-2015-matching-obs.png";

    if(ele[1].checked)
        document.getElementById("1993-vs-2015_plot").src = "images/figures/wid-top-1-1993-vs-2015-matching-obs.png";

  } else {

        if(ele[0].checked)
        document.getElementById("1993-vs-2015_plot").src = "images/figures/pip-gini-1993-vs-2015-all-obs.png";

    if(ele[1].checked)
        document.getElementById("1993-vs-2015_plot").src = "images/figures/wid-top-1-1993-vs-2015-all-obs.png";
  }
}

</script>
</html>

```{python}
output_dict['all_sources_False']['owid_format']

# prepped_data_dict['prepped_data']

# id_vars_duplicates = df[df.duplicated(subset=['source_var', 'Entity'])]

# id_vars_duplicates
```

```{python}
output_dict['all_sources_False']['prepped_data']

- Averages + pop coverage

- Country changes (also see which countries strongly disagree in terms of direction)

## Inequality since 1980

```{python}
# Set-up to compare WID and PIP top 10% share, all observations
reference_vals = [1980, 2018]
max_dist_from_refs = 5

selected_vars = [
        'PIP: Gini',
        'WID: Top 1pc share – pretax'
        ]

output_dict = prep_data_and_change_summary(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      region_col = 'region',
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup
      )

# Store data for OWID upload – prepared further in appendix below
df_owid_1980_2018 = output_dict['owid_format']

fp = 'data_appendix/figures_and_tables/Compare WID and PIP inequality data 1980 vs 2018 (Joe temp).csv'

df_owid_1980_2018.to_csv(fp, index=False)

#### Figures X-Y

```{=html}

<html>
<body>

<input type="radio" name="1980_2018_radio" value="PIP" checked="checked" onclick="chartSwitchFunction1980_2018()">&nbsp;PIP: Gini coefficient
&nbsp;
&nbsp;
<input type="radio" name="1980_2018_radio" value="WID" onclick="chartSwitchFunction1980_2018()">&nbsp;WID: Top 1% share

<br>
<label for="obsCheck">Show only observations available in both datasets:</label> 
<input type="checkbox" id="obsCheck" onclick="chartSwitchFunction1980_2018()">

<iframe id="1980_2018_plot" src="https://ourworldindata.org/grapher/pip-gini-1980-vs-2018-all-obs" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>;

</body>
<script>
function chartSwitchFunction1980_2018() {

  var ele = document.getElementsByName('1980_2018_radio');

  var checkBox = document.getElementById("obsCheck");

  if (checkBox.checked == true){
    if(ele[0].checked)
        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/pip-gini-1980-vs-2018-matching-obs";

    if(ele[1].checked)
        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/wid-top-1-1980-vs-2018-matching-obs";

  } else {

        if(ele[0].checked)
        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/pip-gini-1980-vs-2018-all-obs";

    if(ele[1].checked)
        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/wid-top-1-1980-vs-2018-all-obs";
  }
}

</script>
</html>

# 3. Aligning indices

- Do a 1% vs 1% comparison

- Do a Gini vs Gini comparison

All for 1993 and 1980.

# Bridging the gap for particular countries

@@

# Appendix
```{python}

# df_owid = pd.concat([df_owid_1993_2015,df_owid_1980_2018])

```{python}
# Prep data to upload to OWID for charts

# owid_upload_year_values = {}
# owid_upload_change = {}

# reference_vals = [1993, 2015]
# max_dist_from_refs = 3

# for all_metrics_requirement in [False, True]:

#     prepped_data = prep_data(
#           source_dfs = source_dfs,
#           df_pop = df_pop,
#           df_regions = df_regions,
#           region_col = 'region',
#           selected_vars = selected_vars,
#           reference_vals = reference_vals,
#           max_dist_from_refs = max_dist_from_refs,
#           min_dist_between = 1,
#           all_metrics_requirement = all_metrics_requirement,
#           reference_col = 'Year',
#           group_by_col = 'Entity',
#           tolerance_lookup = tolerance_lookup,
#           outlier_cut_off_upper = None
#           )

#     keep_cols1 = [
#         'source_var', 
#         'Entity',
#         f'Year{reference_vals[0]}', 
#         f'value{reference_vals[0]}', 
#         f'Year{reference_vals[1]}', 
#         f'value{reference_vals[1]}',
#         'region']

#     keep_cols2 = [
#         'source_var', 
#         'Entity',
#         f'Year{reference_vals[0]}', 
#         f'value{reference_vals[0]}', 
#         f'Year{reference_vals[1]}', 
#         f'value{reference_vals[1]}',
#         'region']

#     year_values = prepped_data[keep_cols1]

#     year_values = year_values.pivot_table(
#         index=['Entity', 'region'], 
#         columns='source_var',
#         values=[
#             f'Year{reference_vals[0]}', 
#             f'value{reference_vals[0]}', 
#             f'Year{reference_vals[1]}', 
#             f'value{reference_vals[1]}'
#         ]).reset_index()

#     year_values.columns = ['_'.join([str(col) for col in cols]) for cols in year_values.columns]

#     year_values = year_values.rename(columns={
#         "Entity_": "Entity", 
#         "region_": "region"})

#     year_values = pd.melt(
#         year_values, 
#         id_vars=['Entity', 'region'], 
#         value_vars=[
#             f'Year{reference_vals[0]}_{selected_vars[0]}',
#             f'Year{reference_vals[1]}_{selected_vars[0]}',
#             f'value{reference_vals[0]}_{selected_vars[0]}',
#             f'value{reference_vals[1]}_{selected_vars[0]}',
#             f'Year{reference_vals[0]}_{selected_vars[1]}',
#             f'Year{reference_vals[1]}_{selected_vars[1]}',
#             f'value{reference_vals[0]}_{selected_vars[1]}',
#             f'value{reference_vals[1]}_{selected_vars[1]}'
#         ])

#     year_values[['var', 'source_var']] = year_values['variable'].str.split('_', expand=True)
#     # Extract year and group information from variable column
#     year_values[['var', 'reference']] = year_values['var'].str.extract(r'(\D+)(\d+)$')
#     # owid_format['reference'] = owid_format['reference'].astype(int)

#     #Reshape from long to wide
#     year_values=pd.pivot(year_values, index=['Entity', 'region', 'reference' ], columns = ['var','source_var'] ,values = 'value').reset_index()

#     owid_upload_year_values[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObs{all_metrics_requirement}'] = year_values

    # change_values = prepped_data[keep_cols2]

    # owid_upload_change[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObs{all_metrics_requirement}'] = change_values

# df_owid_year_values = pd.merge(
#     owid_upload_year_values[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObsFalse'],
#     owid_upload_year_values[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObsTrue'],
#     on=['Entity', reference_col, region_col], how='outer', suffixes=(' – all observations', ' – only matching observations'))

# df_owid = owid_upload_year_values[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObsFalse']
# # df_owid = pd.merge(df_owid_1993_2015,df_owid_1980_2018, on=[group_by_col, reference_col, region_col], how='outer', suffixes=(' – all observations', ' – only matching observations'))

# # df_owid = pd.concat([df_owid_1993_2015,df_owid_1980_2018])

# fp = 'data_appendix/figures_and_tables/Compare WID and PIP inequality data (Joe temp).csv'

# df_owid.to_csv(fp, index=False)

```{python}
#  # Reshape the data for OWID upload

#     year_values = pd.melt(prepped_data, id_vars=['source_var', 'Entity', 'region' ], value_vars=[f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', f'value{reference_vals[0]}', f'value{reference_vals[1]}'])

#     # Extract year and group information from variable column
#     year_values[['var', 'reference']] = year_values['variable'].str.extract(r'(\D+)(\d+)$')

#     year_values['reference'] = year_values['reference'].astype(int)

#     #Reshape from long to wide
#     year_values=pd.pivot(year_values, index=['source_var', 'Entity', 'region', 'reference' ], columns = 'var' ,values = 'value').reset_index()

#     # Drop the original reference column
#     year_values = year_values.drop('reference', axis=1)

#     # Pivot the melted dataframe wider – one value column per source_var
#     owid_format = owid_format.pivot_table(index=[group_by_col, region_col, reference_col], columns='source_var', values='value').reset_index()

#     summaries_dict['owid_format'] = owid_format

#     output_dict[f'all_sources_{all_metrics_requirement}'] = summaries_dict

#     # Merge the all_sources- True and False data for OWID format
#     owid_format = pd.merge(output_dict['all_sources_False']['owid_format'], output_dict['all_sources_True']['owid_format'], on=[group_by_col, reference_col, region_col], how='outer', suffixes=(' – all observations', ' – only matching observations'))

#     owid_format = owid_format.rename(columns={group_by_col: "country", reference_col: "year"})

#     # Reorder columns
#     cols = list(owid_format.columns) # Get the list of column names
#     cols.remove('country') # Remove col1 from the list
#     cols.remove('year') # Remove col2 from the list
#     cols = ['country', 'year'] + cols # Add col1 and col2 to the front of the list
#     owid_format = owid_format[cols] # Reorder columns in the DataFrame

#     owid_format['year'] = owid_format['year'].astype(int) # Year needs to be integer for OWID

## Data coverage

## PIP Gini vs WID Top 1% share

### 1993-2015
